### 2015-12-22 article.rmd
Added RMarkdown article format.

---
title: "article"
author: "Emanuel Calvo"
date: "December 22, 2015"
output: html_document
runtime: shiny
---

```{r init, message = FALSE, echo=FALSE}
library(shiny)
require(RMySQL)
source('~/git/fts_article/load.R')

# Yeah, only for development environements
all_cons <- dbListConnections(MySQL())
for(con in all_cons) dbDisconnect(con)

# You do not want clear passwords, but is a sandbox.
con <- dbConnect(MySQL(), 
                 user="msandbox", password="msandbox", port=14901 
                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock", 
                 host="127.0.0.1")
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately.  To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

> You only need to have a MySQL sandbox with the 5.7 version.

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r, echo=FALSE}
        ui <- shinyUI(fluidPage(
                mainPanel(textInput("filtering", "FTS Filter:", ""),
                          submitButton("Plot!"),
                          plotOutput("plot")
                          )
        ))
        server <- function(input, output, session) {
                output$plot<-renderPlot({
                        f<-function(x) {eval(parse(text=input$filtering))}
                        x<-1:10
                        plot(x,f(x))
                }                        
                )              
        }

shinyApp(ui = ui, server = server)
```

Template

```{r}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),

  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")

  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

### 2015-12-23 load.R
Added by line inserts on the book content

@@ -21,23 +21,45 @@ con <- dbConnect(MySQL(),
 ## check all connections have been closed
 # dbListConnections(MySQL())
 
-fileConn<-file("output.txt")
+loadfile <- 'output.txt'
+loadfileline <- 'outputLine.txt'
 
+fileConn<-file(loadfile)
+fileConnByLine<-file(loadfileline)
+
+
+# Raw content of the book by paragraph splitting 
 rawContent <- paste(readLines("pg1232.txt"), collapse = "\n")
 rawContent <- unlist(strsplit(rawContent, "\n[ \t\n]*\n"))
 
+# Raw content by line splitting
+rawContentByLine <- readLines("pg1232.txt")
+rawContentByLine <- unlist(strsplit(rawContentByLine, "\n[ \t\n]*\n"))
+
+
 # Just cleaning for test
 dbGetQuery(con,"DROP TABLE IF EXISTS bookContent;")
+dbGetQuery(con,"DROP TABLE IF EXISTS bookContentByLine;")
+
 
 # Limitation. Using FTS_DOC_ID 
 dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContent
                  (
-                 FTS_DOC_ID BIGINT UNSIGNED AUTO_INCREMENT NOT NULL PRIMARY KEY, 
-                 bookid bigint, content text); ")
+                   FTS_DOC_ID BIGINT UNSIGNED AUTO_INCREMENT PRIMARY KEY, 
+                   bookid bigint, 
+                   content text
+                 ); ")
+
+dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
+                 (
+                   FTS_DOC_ID BIGINT UNSIGNED AUTO_INCREMENT PRIMARY KEY, 
+                   bookid bigint, 
+                   content text
+                 ); ")
 
 
 # This is not the best way to do so. Instead, you want to load into file. IDIOT.
-inserParagraph <- function (bookid, content) {
+insertParagraph <- function (bookid, content) {
   sql <- sprintf("insert into books_content
                (book_id, paragraph)
                values (%d, '%s');",
@@ -46,24 +68,42 @@ inserParagraph <- function (bookid, content) {
 
 }
 
+# bookid = 1 (no other books on this test at the moment)
 tableContent <- data.frame(gsub("\\n", "", rawContent), stringAsFactors=FALSE)
 tableContent <- cbind(1, tableContent )
[…]

### 2015-12-24 article.rmd
Added new code and dynamic shiny page with query against the tables.

@@
                  host="127.0.0.1")
+
 ```
@@
 
-```{r, echo=FALSE}
+```{r, echo=TRUE}
         ui <- shinyUI(fluidPage(
@@
                 output$plot<-renderPlot({
-                        f<-function(x) {eval(parse(text=input$filtering))}
+                        f<-function(x) {as.character(input$filtering)}
+                        sql <- paste("SELECT source, sum(scoreBoolean), sum(scoreQE),
+        max(scoreBoolean),max(scoreQE),
+        avg(scoreBoolean), avg(scoreQE),
+        avg(lengthContent),
+        count(source)
+FROM
+(
+SELECT \"BC\" as source, FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContent BC
+WHERE
+  match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+UNION
+SELECT \"BL\" as source, FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContentByLine BL
+WHERE
+  match(BL.content) against (\"", f(x) ,"\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+) unionTable
+GROUP BY source
+;" , sep = "")
+                        rs <- dbGetQuery(con, sql)
+                        rs
                         x<-1:10
-                        plot(x,f(x))
+                        # plot(x,f(x))
                 }                        

### 2015-12-24 article.rmd
Functional, added pairs just for testing.

--- 2015-12-22 article.rmd
@@
                 output$plot<-renderPlot({
-                        f<-function(x) {eval(parse(text=input$filtering))}
-                        x<-1:10
-                        plot(x,f(x))
+                        f<-function(x) {as.character(input$filtering)}
+
+                        sql <- paste("SELECT sourceTable, sum(scoreBoolean), sum(scoreQE),
+        max(scoreBoolean),max(scoreQE),
+        avg(scoreBoolean), avg(scoreQE),
+        avg(lengthContent),
+        avg(scoreQE) / count(sourceTable) as weight,
+        count(sourceTable)
+FROM
+(
+SELECT \"BC\" as sourceTable, FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContent BC
+WHERE
+  match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+UNION
+SELECT \"BL\" as sourceTable, FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContentByLine BL
+WHERE
+  match(BL.content) against (\"", f(x) ,"\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+) unionTable
+GROUP BY sourceTable
+;" , sep = "")
+                        rs <- dbGetQuery(con, sql)
+                        #rs
+                        #x<-1:10
+
+                        #pairs(rs$sourceTable + rs$weight, data = rs)
+                        pairs(rs[,2:10])
+                        # plot(x,f(x))
                 }                        

### 2015-12-24 article.rmd
Added new entries and plots

--- 2015-12-22 article.rmd
@@
 require(RMySQL)
+library(ggplot2)
+
 source('~/git/fts_article/load.R')
@@
                 output$plot<-renderPlot({
-                        f<-function(x) {eval(parse(text=input$filtering))}
-                        x<-1:10
-                        plot(x,f(x))
+                        f<-function(x) {as.character(input$filtering)}
+
+                        sql <- paste("SELECT sourceTable, 
+        sum(scoreBoolean) as sumBoolean, 
+        sum(scoreQE) as sumQE,
+        max(scoreBoolean) as maxBoolean,
+        max(scoreQE) as maxBoolean,
+        avg(scoreBoolean) as avgBoolean, 
+        avg(scoreQE) as avgQE,
+        avg(lengthContent) avgLength,
+        avg(scoreQE) / count(sourceTable) as weight,
+        count(sourceTable) as countOccurrences
+FROM
+(
+SELECT \"BC\" as sourceTable, FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContent BC
+WHERE
+  match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+UNION
+SELECT \"BL\" as sourceTable, FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContentByLine BL
+WHERE
+  match(BL.content) against (\"", f(x) ,"\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+) unionTable
+GROUP BY sourceTable
+;" , sep = "")
+                        rs <- dbGetQuery(con, sql)
+                        #rs
+                        #x<-1:10
+
+                        #pairs(rs$sourceTable + rs$weight, data = rs)
+                        pairs(rs[,2:10])
+                        # plot(x,f(x))
                 }                        

### 2015-12-24 plots.R

@@ -1,0 +1,78 @@
+require(RMySQL)
+library(ggplot2)
+require(caret)
+
+
+source('~/git/fts_article/load.R')
+
+# Yeah, only for development environements
+all_cons <- dbListConnections(MySQL())
+for(con in all_cons) dbDisconnect(con)
+
+
+# You do not want clear passwords, but is a sandbox.
+con <- dbConnect(MySQL(), 
+                 user="msandbox", password="msandbox", port=14901 
+                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock", 
+                 host="127.0.0.1")
+
+word <- "country"
+
+#f<-function(x) {as.character(input$filtering)}
+
+f <- function(x) {as.character(word)}
+
+sql <- paste("SELECT sourceTable, sum(scoreBoolean) as sumBoolean, 
+        sum(scoreQE) as sumQE,
+        max(scoreBoolean) as maxBoolean,
+        max(scoreQE) as maxBoolean,
+        avg(scoreBoolean) as avgBoolean, 
+        avg(scoreQE) as avgQE,
+        avg(lengthContent) avgLength,
+        avg(scoreQE) / count(sourceTable) as weight,
+        count(sourceTable) as countOccurrences
+             FROM
+             (
+             SELECT \"BC\" as sourceTable, FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContent BC
+WHERE
+  match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+UNION
+SELECT \"BL\" as sourceTable, FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
+       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       length(content) as lengthContent
+FROM
+        bookContentByLine BL
+WHERE
[…]

### 2015-12-24 article.rmd
http://stackoverflow.com/questions/19706320/create-a-data-frame-using-text-input-in-shiny

@@
 
-## Inputs and Outputs
+## Moto of the article
+
+A couple of days ago one of our customers came up with a question regarding FTS over InnoDB engine. Althuogh the question
+is not entirely related with most of the content written here, I came up with the conclusion that FTS sometimes is a misunderstood feature.
+
+Both official documentation and articles are propagated over internet, however sometimes most of them lacks to answer the
+main questions and test cases. In order to follow a reproducible article, I used public available data sources (www.gutenberg.org).
+
+In order to show the effects off the field sizes on the ranking algorithms, you will see two main tables (bookContent and 
+bookContentByLine) which contains the books parsed by paragraph and by line respectively. You'll see the noise generated
+by the `QUERY EXPANSION` algorithm when phrases are too large, and that's why you see both tables.
+
+The current article has been developed using Shiny/R in order to allow you to see the effects of the algorithms.
+
+Here is an example of how ranks differ among algorithms and field sizes using the word 'country':
+
+```
+set global innodb_ft_aux_table = 'test/bookContentByLine';
+
+SELECT  content, group_concat(it.POSITION) as positions, 
+        round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank, 
+        round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, 
+        length(content) as len 
+      FROM bookContentByLine bl join information_schema.INNODB_FT_INDEX_TABLE it 
+          ON (bl.FTS_DOC_ID = it.DOC_ID)  
+      WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) 
+          AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY 3 DESC LIMIT 10 ;
+
++---------------------------------------------------------------------------+-----------+--------+-------------+------+
+| content                                                                   | positions | QERank | BooleanRank | len  |
++---------------------------------------------------------------------------+-----------+--------+-------------+------+
+| "country may be ennobled, and under its auspices may be verified that"    | 1         |     54 |           4 |   70 |
+| "country, under a most excellent president, wherein all cities had their" | 1         |     53 |           4 |   73 |
+| "Again, the prince who holds a country differing in the above respects"   | 31        |     51 |           4 |   71 |
+| "country districts. As Lucca had five gates, he divided his own country"  | 1,63      |     51 |           8 |   72 |
+| "inflicting immense damage upon the country. Whereupon the Florentines"   | 36        |     50 |           4 |   71 |
+| "The cities of Germany are absolutely free, they own but little country"  | 64        |     49 |           4 |   72 |
+| "citizens of Fermo, to whom the slavery of their country was dearer than" | 49        |     48 |           4 |   73 |
+| "is useful in two ways. Firstly, he learns to know his country, and"      | 55        |     46 |           4 |   68 |
+| "agree that he should retain any lordship over the country. Because the"  | 51        |     45 |           4 |   72 |
+| "his own parts of the country, according to the authority he had assumed" | 22        |     43 |           4 |   73 |
++---------------------------------------------------------------------------+-----------+--------+-------------+------+
+
+node1 [localhost] {msandbox} (test) > set global innodb_ft_aux_table = 'test/bookContent';
+Query OK, 0 rows affected (0,00 sec)
+
+node1 [localhost] {msandbox} (test) > SELECT FTS_DOC_ID, group_concat(it.POSITION) as positions, round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank, round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, length(content) as len FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY QERank DESC LIMIT 10 ;
++------------+------------------+--------+-------------+------+
+| FTS_DOC_ID | positions        | QERank | BooleanRank | len  |
++------------+------------------+--------+-------------+------+
+|        757 | 6858             |   2804 |           1 | 8101 |
+|        735 | 2842             |    925 |           1 | 3133 |
+|        157 | 747              |    698 |           1 | 2385 |
+|        733 | 265,1342,816,750 |    569 |           2 | 1765 |
+|        739 | 1171             |    540 |           1 | 1659 |
+|        249 | 197,739          |    435 |           1 | 1368 |
+|        279 | 664              |    428 |           1 | 1365 |
+|        411 | 468,410          |    396 |           2 | 1170 |
+|        155 | 31,495,158,525   |    363 |           5 | 1378 |
+|        327 | 60               |    362 |           1 | 1372 |
++------------+------------------+--------+-------------+------+
+10 rows in set (4,09 sec)
+```
+
+```{r, message=FALSE, echo=FALSE}
+sql <- "SELECT avg(length(content)) as len from bookContent"
+rs <- dbGetQuery(con,sql)
+```
+
+Following stats are from the bookContent table (which has an average length of `r rs$len` characters )
+
+```
+node1 [localhost] {msandbox} (test) > select * from information_schema.INNODB_FT_INDEX_TABLE WHERE DOC_ID = 733 AND lower(WORD) like '%country%';
++------------------+--------------+-------------+-----------+--------+----------+
+| WORD             | FIRST_DOC_ID | LAST_DOC_ID | DOC_COUNT | DOC_ID | POSITION |
++------------------+--------------+-------------+-----------+--------+----------+
+| country          |          149 |         787 |        28 |    733 |      265 |
+| country          |          149 |         787 |        28 |    733 |     1342 |
+| countrydistricts |          733 |         733 |         1 |    733 |      816 |
+| thecountry       |          249 |         733 |         2 |    733 |      750 |
++------------------+--------------+-------------+-----------+--------+----------+
+4 rows in set (0,08 sec)
+
+node1 [localhost] {msandbox} (test) > select * from information_schema.INNODB_FT_INDEX_TABLE WHERE DOC_ID = 155 AND lower(WORD) like '%country%';
++---------+--------------+-------------+-----------+--------+----------+
+| WORD    | FIRST_DOC_ID | LAST_DOC_ID | DOC_COUNT | DOC_ID | POSITION |
++---------+--------------+-------------+-----------+--------+----------+
+| country |          149 |         787 |        28 |    155 |       31 |
+| country |          149 |         787 |        28 |    155 |      495 |
+| country |          149 |         787 |        28 |    155 |      158 |
+| country |          149 |         787 |        28 |    155 |      525 |
++---------+--------------+-------------+-----------+--------+----------+
+4 rows in set (0,09 sec)
+
+```
+
+> In the example shown before the is no intention to compare ranks score as they are based in different algorithms. 
+> The idea there is to show that QUERY EXPANSION can have non desire results in some cases due to its mechanism.
+
+## Playing with searches
 
@@
                 mainPanel(textInput("filtering", "FTS Filter:", ""),
-                          submitButton("Plot!"),
-                          plotOutput("plot")
+                          submitButton("Stats!"),
+                          #plotOutput("plot")
+                          uiOutput('table')
                           )
@@
 
+                        output$table <- renderTable({
+                            if (is.null(Data())) {return()}
+                            print(Data()$rs)
+                          }, 'include.rownames' = FALSE
+                           , 'include.colnames' = TRUE
+                           , 'sanitize.text.function' = function(x){x}
+                        )
+
                         #pairs(rs$sourceTable + rs$weight, data = rs)
-                        pairs(rs[,2:10])
+                        #pairs(rs[,2:10])
                         # plot(x,f(x))

### 2015-12-24 plots.R

@@ -16,7 +16,7 @@ con <- dbConnect(MySQL(),
                  ,dbname="test", socket="/tmp/mysql_sandbox14901.sock", 
                  host="127.0.0.1")
 
-word <- "country"
+word <- "find"
 
 #f<-function(x) {as.character(input$filtering)}
 
@@ -54,25 +54,32 @@ GROUP BY sourceTable
 rs <- dbGetQuery(con, sql)
 
 
+QE <- "WITH QUERY EXPANSION"
+# QE <- ""
+
 sql <- paste("
 SELECT \"BC\" as sourceTable, FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
-       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE ",QE,") scoreQE,
        length(content) as lengthContent
 FROM
         bookContent BC
 WHERE
-  match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+  match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE ",QE,")
 UNION
 SELECT \"BL\" as sourceTable, FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
-       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE ",QE,") scoreQE,
        length(content) as lengthContent
 FROM
         bookContentByLine BL
 WHERE
-  match(BL.content) against (\"", f(x) ,"\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+  match(BL.content) against (\"", f(x) ,"\" IN NATURAL LANGUAGE MODE ",QE,")
 ")
 
 rs <- dbGetQuery(con, sql)
 
+pl1 <- qplot(scoreBoolean,scoreQE, data = rs, fill = factor(sourceTable), geom = c("boxplot") )
+
[…]

### 2015-12-25 article.rmd
data table output working

@@
 ---
-title: "article"
+title: "Full text search hack with Shiny/R"
 author: "Emanuel Calvo"
@@
 ```{r init, message = FALSE, echo=FALSE}
+options(width = 9999)
+knitr::opts_chunk$set(width=9999)
+
+library(knitr)
 library(shiny)
@@
 is not entirely related with most of the content written here, I came up with the conclusion that FTS sometimes is a misunderstood feature.
+
+The point of this article will be clarify when each algorithm fits in each situation.
 
@@
 
+As we can observ on the previous examples, the query extension algorithm has some noisy produced by the double query
+generated in order to join  the results.
+
 ```{r, message=FALSE, echo=FALSE}
@@
 
-Following stats are from the bookContent table (which has an average length of `r rs$len` characters )
+Following stats are from the bookContent table (which has an average length of `r rs$len` characters ).
 
@@
 
-You can embed Shiny inputs and outputs in your document.
-Outputs are automatically updated whenever inputs change.
-This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function.
-The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.
+The following form will allow you to search across the stats over the tables
+of the filtering provided dynamically.
 
-```{r, echo=TRUE}
+Keep in mind that the numbers are not meaningful as we can't compare rankings
+between `query extension` and `boolean`.
+
+```{r, echo=FALSE}
         ui <- shinyUI(fluidPage(
-                mainPanel(textInput("filtering", "FTS Filter:", ""),
+                mainPanel(textInput("filtering", "FTS Filter:", "default"),
                           submitButton("Stats!"),
-                          #plotOutput("plot")
                           uiOutput('table')
@@
         ))
+
         server <- function(input, output, session) {
-                output$plot<-renderPlot({
+
+                Data = reactive({
                   f<-function(x) {as.character(input$filtering)}
@@
                               (
-SELECT \"BC\" as sourceTable, FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
-       match(BC.content) against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+                              SELECT \"byParagraph\" as sourceTable, 
+                                    FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
+                                     match(BC.content) 
+                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                      length(content) as lengthContent
-FROM
-        bookContent BC
+                              FROM    bookContent BC
                               WHERE
@@
                               UNION
-SELECT \"BL\" as sourceTable, FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
-       match(BL.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+                              SELECT \"byLine\" as sourceTable, 
+                                      FTS_DOC_ID, match(BL.content) against (\"",f(x)," \" IN BOOLEAN MODE) as scoreBoolean,
+                                     match(BL.content) 
+                                      against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
                                      length(content) as lengthContent
@@
                               ;" , sep = "")
+
                       rs <- dbGetQuery(con, sql)
-                        #rs
-                        #x<-1:10
+                      return(rs)
 
-                        output$table <- renderTable({
-                            if (is.null(Data())) {return()}
-                            print(Data()$rs)
-                          }, 'include.rownames' = FALSE
-                           , 'include.colnames' = TRUE
-                           , 'sanitize.text.function' = function(x){x}
-                        )
+                })
 
-                        #pairs(rs$sourceTable + rs$weight, data = rs)
-                        #pairs(rs[,2:10])
-                        # plot(x,f(x))
-                }                        
-                )              
+                        output$table <- renderTable({Data()})
+
         }
@@
 shinyApp(ui = ui, server = server)
+
 ```

### 2015-12-26 article.rmd
Starting to add more intelligence when downloading files. Needs fix.

@@
 As we can observ on the previous examples, the query extension algorithm has some noisy produced by the double query
-generated in order to join  the results.
+generated in order to join  the results. Also, it won't matter if you have the exact match twice, as the algorithm does not
+focus on the exact matches, instead it will relate the occurrences on the first search (where the match happens) and the
+sencond (where the algorithm searches for the words that appear in the same phrase than the exact match). That means 
+that it is possible to have a high rank without have any matching occurrence.
+
+This kind of searches is useful if you want to find entries with words that appear frequently within your search.
 

### 2015-12-26 load.R

@@ -1,10 +1,26 @@
 require(RMySQL)
 require(utils)
+library(optparse) # https://github.com/trevorld/optparse
+                  # devtools::install_github("trevorld/optparse")
+
+randnums <- 5
+set.seed(1)
 
 setwd("~/git/fts_article/")
 
-# Pretty basic check, but useful.
-if(!file.exists('pg1232.txt')) download.file('http://www.gutenberg.org/cache/epub/1232/pg1232.txt', 'pg1232.txt')
+
+parser <- OptionParser()
+#parser <- add_option(parser, c("-v", "--verbose"), action="store_true",
+#                                        default=TRUE, help="Print extra output [default]")
+#parser <- add_option(parser, c("-q", "--quietly"), action="store_false",
+#                                            dest="verbose", help="Print little output")
+parser <- add_option(parser, c("-b", "--books"), type="integer", default=5,
+                                        help="Number of books to download [default %default]",
+                                        metavar="randnums")
+#parse_args(parser, args = c("--quietly", "--count=15"))
+parse_args(parser)
+
+# Probably adding more stuff here
 
 # Horrible, but due to an issue when working with debugging in RStudio, some
 # connections halt
@@ -18,30 +34,10 @@ con <- dbConnect(MySQL(),
                  host="127.0.0.1")
 # on.exit(dbDisconnect(con))
 
-## check all connections have been closed
-# dbListConnections(MySQL())
-
-loadfile <- 'output.txt'
-loadfileline <- 'outputLine.txt'
-
-fileConn<-file(loadfile)
-fileConnByLine<-file(loadfileline)
-
-
-# Raw content of the book by paragraph splitting 
-rawContent <- paste(readLines("pg1232.txt"), collapse = "\n")
-rawContent <- unlist(strsplit(rawContent, "\n[ \t\n]*\n"))
-
-# Raw content by line splitting
-rawContentByLine <- readLines("pg1232.txt")
-rawContentByLine <- unlist(strsplit(rawContentByLine, "\n[ \t\n]*\n"))
-
 
 # Just cleaning for test
 dbGetQuery(con,"DROP TABLE IF EXISTS bookContent;")
 dbGetQuery(con,"DROP TABLE IF EXISTS bookContentByLine;")
-
-
[…]

### 2015-12-26 load.R
Fixed

@@ -18,7 +18,7 @@ parser <- add_option(parser, c("-b", "--books"), type="integer", default=5,
                                         help="Number of books to download [default %default]",
                                         metavar="randnums")
 #parse_args(parser, args = c("--quietly", "--count=15"))
-parse_args(parser)
+opts <- parse_args(parser)
 
 # Probably adding more stuff here
 
@@ -59,12 +59,12 @@ randids <- abs(round(rnorm(randnums)*1000))
 # How many downloaded books?
 rest <- randnums - length(dir(pattern = "pg.*txt"))
 
-for (bi in 1:randnums) {
+for (bi in 1:opts$randnums) {
   tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 
                         ,randids[bi], 
                         '/pg',randids[bi],'.txt', sep = ""
                         )
-  if(!file.exists(tempFileName)) {next} 
+  if(file.exists(tempFileName)) {next} 
   download.file(tempfileName, paste('pg',randids[bi],'.txt'), sep = "")
 }
 

### 2015-12-26 load.R
gutenberg doesn't allow to download that much. Downloading a couple by hand, controlled the amount of books using --books parameter

@@ -3,7 +3,7 @@ require(utils)
 library(optparse) # https://github.com/trevorld/optparse
                   # devtools::install_github("trevorld/optparse")
 
-randnums <- 5
+#randnums <- 5
 set.seed(1)
 
 setwd("~/git/fts_article/")
@@ -18,7 +18,8 @@ parser <- add_option(parser, c("-b", "--books"), type="integer", default=5,
                                         help="Number of books to download [default %default]",
                                         metavar="randnums")
 #parse_args(parser, args = c("--quietly", "--count=15"))
-parse_args(parser)
+opts <- parse_args(parser)
+
 
 # Probably adding more stuff here
 
@@ -54,26 +55,23 @@ dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
                  ); ")
 
 #vector of ids for book downloading
-randids <- abs(round(rnorm(randnums)*1000))
+#randids <- abs(round(rnorm(opts$books)*1000))
 
 # How many downloaded books?
-rest <- randnums - length(dir(pattern = "pg.*txt"))
+#rest <- randnums - length(dir(pattern = "pg.*txt"))
 
-for (bi in 1:randnums) {
-  tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 
-                        ,randids[bi], 
-                        '/pg',randids[bi],'.txt', sep = ""
-                        )
-  if(!file.exists(tempFileName)) {next} 
-  download.file(tempfileName, paste('pg',randids[bi],'.txt'), sep = "")
-}
+#for (bi in 1:opts$books) {
+#  tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 
+#                        ,randids[bi], 
+#                        '/pg',randids[bi],'.txt', sep = ""
+#                        )
+#  if(file.exists(tempFileName)) {next} 
+#  download.file(tempFileName, paste('pg',randids[bi],'.txt', sep = ""))
+#}
 
 
 # Pretty basic check, but useful.
 #if(!file.exists(paste('pg', randids ,'.txt'))) download.file('http://www.gutenberg.org/cache/epub/1232/pg1232.txt', 'pg1232.txt')
-
-
-
 
 ## check all connections have been closed
 # dbListConnections(MySQL())
@@ -87,7 +85,7 @@ laodFilesLine <- paste('ouputLine' , rids, sep = "")
[…]

### 2015-12-26 load.R
Fixing merge

@@ -61,7 +61,7 @@ dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
 #randids <- abs(round(rnorm(opts$books)*1000))
 
 # How many downloaded books?
-<<<<<<< HEAD
+
 #rest <- randnums - length(dir(pattern = "pg.*txt"))
 
 #for (bi in 1:opts$books) {
@@ -72,18 +72,18 @@ dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
 #  if(file.exists(tempFileName)) {next} 
 #  download.file(tempFileName, paste('pg',randids[bi],'.txt', sep = ""))
 #}
-=======
+
 rest <- randnums - length(dir(pattern = "pg.*txt"))
 
-for (bi in 1:opts$randnums) {
-  tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 
-                        ,randids[bi], 
-                        '/pg',randids[bi],'.txt', sep = ""
-                        )
-  if(file.exists(tempFileName)) {next} 
-  download.file(tempfileName, paste('pg',randids[bi],'.txt'), sep = "")
-}
->>>>>>> 487fd798b3e2aa97a33f40bdf9e53d40d979d1a9
+#for (bi in 1:opts$randnums) {
+#  tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 
+#                        ,randids[bi], 
+#                        '/pg',randids[bi],'.txt', sep = ""
+#                        )
+#  if(file.exists(tempFileName)) {next} 
+#  download.file(tempfileName, paste('pg',randids[bi],'.txt'), sep = "")
+#}
+
 
 
 # Pretty basic check, but useful.

### 2015-12-26 load.R
Changed radically the way to insert files into the database (files already downloaded).

@@ -1,5 +1,6 @@
 require(RMySQL)
 require(utils)
+library(stringr)
 library(optparse) # https://github.com/trevorld/optparse
                   # devtools::install_github("trevorld/optparse")
 
@@ -8,7 +9,7 @@ set.seed(1)
 
 setwd("~/git/fts_article/")
 
-
+#parsing arguments for future purposes
 parser <- OptionParser()
 #parser <- add_option(parser, c("-v", "--verbose"), action="store_true",
 #                                        default=TRUE, help="Print extra output [default]")
@@ -19,10 +20,7 @@ parser <- add_option(parser, c("-b", "--books"), type="integer", default=5,
                                         metavar="randnums")
 #parse_args(parser, args = c("--quietly", "--count=15"))
 opts <- parse_args(parser)
-<<<<<<< HEAD
 
-=======
->>>>>>> 487fd798b3e2aa97a33f40bdf9e53d40d979d1a9
 
 # Probably adding more stuff here
 
@@ -73,7 +71,7 @@ dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
 #  download.file(tempFileName, paste('pg',randids[bi],'.txt', sep = ""))
 #}
 
-rest <- randnums - length(dir(pattern = "pg.*txt"))
+rest <- opts$books - length(dir(path = "./books",pattern = "pg.*txt"))
 
 #for (bi in 1:opts$randnums) {
 #  tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 
@@ -92,22 +90,28 @@ rest <- randnums - length(dir(pattern = "pg.*txt"))
 ## check all connections have been closed
 # dbListConnections(MySQL())
 
-loadfile <- 'output'
-loadfileline <- 'outputLine'
+#loadfile <- 'output'
+#loadfileline <- 'outputLine'
 
-loadFiles <- paste('ouput' ,rids, sep = "")
-laodFilesLine <- paste('ouputLine' , rids, sep = "")
+#loadFiles <- paste('ouput' ,rids, sep = "")
+#loadFilesLine <- paste('ouputLine' , rids, sep = "")
 
-fileConn<-file(loadfile)
-fileConnByLine<-file(loadfileline)
+#fileConn<-file(loadfile)
+#fileConnByLine<-file(loadfileline)
 
-listOfDownloadedFiles <- sort(dir(path = "./books/", pattern = "pg.*txt"))
+listOfDownloadedFiles <- sort(dir(path = "./books", 
[…]

### 2015-12-26 load.R
Fixes.

@@ -16,11 +16,12 @@ parser <- OptionParser()
 #parser <- add_option(parser, c("-q", "--quietly"), action="store_false",
 #                                            dest="verbose", help="Print little output")
 parser <- add_option(parser, c("-b", "--books"), type="integer", default=5,
-                                        help="Number of books to download [default %default]",
-                                        metavar="randnums")
+                                        help="Number of books to download [default %default]")
+                                        #metavar="randnums")
 #parse_args(parser, args = c("--quietly", "--count=15"))
 opts <- parse_args(parser)
 
+numbooks <- opts$books
 
 # Probably adding more stuff here
 
@@ -71,7 +72,7 @@ dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
 #  download.file(tempFileName, paste('pg',randids[bi],'.txt', sep = ""))
 #}
 
-rest <- opts$books - length(dir(path = "./books",pattern = "pg.*txt"))
+rest <- numbooks - length(dir(path = "./books",pattern = "pg.*txt"))
 
 #for (bi in 1:opts$randnums) {
 #  tempFileName <- paste('http://www.gutenberg.org/cache/epub/' 

### 2015-12-26 article.rmd
Fixes.
Fixes

@@
 ---
-title: "Full text search hack with Shiny/R"
+title: "InnoDB Full text Search hacking with Shiny/R"
 author: "Emanuel Calvo"
@@
 ```{r init, message = FALSE, echo=FALSE}
-options(width = 9999)
-knitr::opts_chunk$set(width=9999)
+#options(width = 9999)
+#knitr::opts_chunk$set(width=9999)
 
@@
 
-source('~/git/fts_article/load.R')
+#Comment if you want to avoid reload all the data
+#source('~/fts_article/load.R')
 
@@
 
+```
+
+As we can observ on the previous examples, the query extension algorithm has some noisy produced by the double query
+generated in order to join  the results. Also, it won't matter if you have the exact match twice, as the algorithm does not
+focus on the exact matches, instead it will relate the occurrences on the first search (where the match happens) and the
+sencond (where the algorithm searches for the words that appear in the same phrase than the exact match). That means 
+that it is possible to have a high rank without have any exact matching occurrence.
+
+The interesting case is the following row, which has 2 exact occurrences and it is not the highest rank using query
+extension. Remember, this is expected.
+
+```
+| "country districts. As Lucca had five gates, he divided his own country"  | 1,63      |     51 |           8 |   72 |
+```
+
+Even worst when using large sentences. In the example bellow you will see the same query, against the table 
+storing by paragraph. The boolean rank shows some of the entries way above others, however the query extension
+locates at the top records that not necessarily has a lot of exact matches.
+
+```
+
 node1 [localhost] {msandbox} (test) > set global innodb_ft_aux_table = 'test/bookContent';
@@
 
-As we can observ on the previous examples, the query extension algorithm has some noisy produced by the double query
-generated in order to join  the results. Also, it won't matter if you have the exact match twice, as the algorithm does not
-focus on the exact matches, instead it will relate the occurrences on the first search (where the match happens) and the
-sencond (where the algorithm searches for the words that appear in the same phrase than the exact match). That means 
-that it is possible to have a high rank without have any matching occurrence.
-
 This kind of searches is useful if you want to find entries with words that appear frequently within your search.
@@
 ```
-node1 [localhost] {msandbox} (test) > select * from information_schema.INNODB_FT_INDEX_TABLE WHERE DOC_ID = 733 AND lower(WORD) like '%country%';
+(test) > select * 
+          from information_schema.INNODB_FT_INDEX_TABLE 
+          WHERE DOC_ID = 733 AND lower(WORD) like '%country%';
 +------------------+--------------+-------------+-----------+--------+----------+
@@
 
-node1 [localhost] {msandbox} (test) > select * from information_schema.INNODB_FT_INDEX_TABLE WHERE DOC_ID = 155 AND lower(WORD) like '%country%';
+(test) > select * 
+          from information_schema.INNODB_FT_INDEX_TABLE 
+          WHERE DOC_ID = 155 AND lower(WORD) like '%country%';
 +---------+--------------+-------------+-----------+--------+----------+
@@
 
+Now that you can see the general stats, we'll see how much noise has been added with the query extension algorithm.
+To do that,  we'll gather all the occurrences using the query extension option and extract only the ones that match 
+with the occurrence.
+
 Template

### 2015-12-27 article.rmd
Added a few more capabilities.

@@
 
-The following form will allow you to search across the stats over the tables
-of the filtering provided dynamically.
+The following form will allow you to search across the stats over the tables of the filtering provided dynamically.
 
-Keep in mind that the numbers are not meaningful as we can't compare rankings
-between `query extension` and `boolean`.
+The aggregations shown on the results are over the results using the selected algorithm. The aggregations over
+the query expansion ranks are only against the results, not all the rows.
 
@@
         ui <- shinyUI(fluidPage(
-                mainPanel(textInput("filtering", "FTS Filter:", "default"),
+                mainPanel(  
+                          selectInput("ftsMode", label = "Algorithm:",
+                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXTENSION"),
+                          textInput("filtering", "FTS Filter:", "default"),
                           submitButton("Stats!"),
@@
                       sql <- paste("SELECT sourceTable, 
-                                      sum(scoreBoolean) as sumBoolean, 
-                                      sum(scoreQE) as sumQE,
                                       max(scoreBoolean) as maxBoolean,
-                                      max(scoreQE) as maxBoolean,
+                                      max(scoreQE) as maxQE,
+                                      max(scoreNL) as maxNL,
                                       avg(scoreBoolean) as avgBoolean, 
                                       avg(scoreQE) as avgQE,
-                                      avg(lengthContent) avgLength,
-                                      avg(scoreQE) / count(sourceTable) as weight,
+                                      avg(scoreNL) as avgNL,
+                                      avg(lengthContent) avgFieldLength,
+                                      -- avg(scoreQE) / count(sourceTable) as weight,
                                       count(sourceTable) as countOccurrences
@@
                               SELECT \"byParagraph\" as sourceTable, 
-                                    FTS_DOC_ID, match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
+                                    FTS_DOC_ID, 
+                                    match(BC.content) against (\"" ,f(x) ," \" IN BOOLEAN MODE) as scoreBoolean,
                                     match(BC.content) 
                                         against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+                                    match(BC.content) 
+                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                      length(content) as lengthContent
@@
                               WHERE
-                                match(BC.content) against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+                                match(BC.content) against (\"" ,f(x)," \" ",input$ftsMode,")
                               UNION
@@
                                       against (\"" ,f(x)," \" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) scoreQE,
+                                    match(BL.content) 
+                                        against (\""  ,f(x), "\" IN NATURAL LANGUAGE MODE) scoreNL,
                                      length(content) as lengthContent
@@
                               WHERE
-                                match(BL.content) against (\"", f(x) ,"\" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
+                                match(BL.content) against (\"", f(x) ,"\" ",input$ftsMode,")
                               ) unionTable
@@
 In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.
+
+### Fine tuning
+
+[Fine tuning](https://dev.mysql.com/doc/refman/5.7/en/fulltext-fine-tuning.html)
+
+### Maintenance
+
+[innodb_optimize_fulltext_only](https://dev.mysql.com/doc/refman/5.7/en/innodb-parameters.html#sysvar_innodb_optimize_fulltext_only)
+
+### Parsers internals
+
+[Writting FTS parser plugins](https://dev.mysql.com/doc/refman/5.7/en/writing-full-text-plugins.html)

### 2015-12-27 article.rmd
Added new sections

@@
 
+## Stats
+
+```{r, echo=FALSE}
+
+sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength 
+            FROM bookContent bp  
+            GROUP BY bookid;"
+rs <- dbGetQuery(con, sql)
+return(rs)
+
+sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength 
+            FROM  bookContentByLine bl 
+            GROUP BY bookid;"
+rs <- dbGetQuery(con, sql)
+return(rs)
+```
+
 ## Playing with searches
@@
 the query expansion ranks are only against the results, not all the rows.
+
+You can start trying `-country +president` in boolean mode.
 
@@
                           submitButton("Stats!"),
+                          # uiOutput('tableStat'),
                           uiOutput('table')
@@
         server <- function(input, output, session) {
+
+                #DataStat = reactive({
+                #  sql <- paste()
+                #  
+                #  rs <- dbGetQuery(con, sql)
+                #  return(rs)
+                #  
+                #})
+
+                #output$tableStat <- renderTable({DataStat()})
 
@@
                                       max(scoreQE) as maxQE,
-                                      max(scoreNL) as maxNL,
+                                      -- max(scoreNL) as maxNL,
                                       avg(scoreBoolean) as avgBoolean, 
                                       avg(scoreQE) as avgQE,
-                                      avg(scoreNL) as avgNL,
+                                      -- avg(scoreNL) as avgNL,
                                       avg(lengthContent) avgFieldLength,

### 2015-12-27 load.R

@@ -179,6 +179,8 @@ sapply(listOfLoadFilesLine, loadFilesLine)
 
 dbGetQuery(con, "CREATE FULLTEXT INDEX ftscontent ON bookContent(content);")
 dbGetQuery(con, "CREATE FULLTEXT INDEX ftscontent ON bookContentByLine(content);")
+dbGetQuery(con, "CREATE INDEX ftsbookid ON bookContent(bookid);")
+dbGetQuery(con, "CREATE INDEX ftsbookid ON bookContentByLine(bookid);")
 
 ### Turning down the music
 dbDisconnect(con)

### 2015-12-27 article.rmd
fixes

@@
 
-As we can observ on the previous examples, the query extension algorithm has some noisy produced by the double query
-generated in order to join  the results. Also, it won't matter if you have the exact match twice, as the algorithm does not
-focus on the exact matches, instead it will relate the occurrences on the first search (where the match happens) and the
-sencond (where the algorithm searches for the words that appear in the same phrase than the exact match). That means 
-that it is possible to have a high rank without have any exact matching occurrence.
+The noise generated by the query expansion is expected and described in the official documentation [here.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-query-expansion.html)
 
@@
 
-Even worst when using large sentences. In the example bellow you will see the same query, against the table 
+This is even worser when using large sentences. In the example bellow you will see the same query, against the table 
 storing by paragraph. The boolean rank shows some of the entries way above others, however the query extension
@@
 
-node1 [localhost] {msandbox} (test) > set global innodb_ft_aux_table = 'test/bookContent';
+(test) > set global innodb_ft_aux_table = 'test/bookContent';
 Query OK, 0 rows affected (0,00 sec)
 
-node1 [localhost] {msandbox} (test) > SELECT FTS_DOC_ID, group_concat(it.POSITION) as positions, round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank, round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, length(content) as len FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY QERank DESC LIMIT 10 ;
+(test) > SELECT FTS_DOC_ID, 
+                group_concat(it.POSITION) as positions, 
+                round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank, 
+                round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, 
+                length(content) as len 
+          FROM bookContent bl join information_schema.INNODB_FT_INDEX_TABLE it ON (bl.FTS_DOC_ID = it.DOC_ID)  
+          WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION) 
+              AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY QERank DESC LIMIT 10 ;
 +------------+------------------+--------+-------------+------+
@@
 +------------+------------------+--------+-------------+------+
-|        757 | 6858             |   2804 |           1 | 8101 |
+|        757 | 6858             |   2804 |           1 | 8101 | <<< This one on the top does only 1 exact match.
 |        735 | 2842             |    925 |           1 | 3133 |
@@
 |        411 | 468,410          |    396 |           2 | 1170 |
-|        155 | 31,495,158,525   |    363 |           5 | 1378 |
+|        155 | 31,495,158,525   |    363 |           5 | 1378 |  <<< This doc contains 5 exact matches!
 |        327 | 60               |    362 |           1 | 1372 |
@@
 
-sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength 
+sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength, count(*) as numEntries 
             FROM bookContent bp  
@@
 
-sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength 
+bookContentStats <- rs
+
+sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength, count(*) as numEntries 
             FROM  bookContentByLine bl 
@@
 return(rs)
+
+bookContentByLineStats <- rs
+
 ```
@@
 
-The aggregations shown on the results are over the results using the selected algorithm. The aggregations over
-the query expansion ranks are only against the results, not all the rows.
+The aggregations shown on the results are over the results using the selected algorithm.
 
-You can start trying `-country +president` in boolean mode.
+You can start trying `+country +(>house <president) -home` in boolean mode. For more information about operator in Innodb's FTS,
+you can read [this.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-boolean.html) Using boolean formulas in NATURAL LANGUAGE MODE,
+does not have the desired effect, however you'll be able to see how it affects the ranking compared with the boolean search.
+
+NOTE: distance operator not yet supported.
 
@@
                           selectInput("ftsMode", label = "Algorithm:",
-                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXTENSION"),
+                                      choices = c("WITH QUERY EXPANSION", "IN BOOLEAN MODE", "IN NATURAL LANGUAGE MODE"), selected = "WITH QUERY EXPANSION"),
                           textInput("filtering", "FTS Filter:", "default"),
@@
 
+## Recommendations
+
+- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
+  size is crucial when using this.
+- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
+  single fields. 
+- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the 
+  FTS functions to avoid double executions. You can do this using somehting like:
+  `match(content,title) against ("first (<second >third)") > 1 `
+  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches. 
+- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
+  maintance on the table.
+- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
+
+---------------------- FROM HERE ONLY TRASH TO BE REMOVED -----------------------------
+
 Template

### 2015-12-27 article.rmd
Added new table output

@@
 author: "Emanuel Calvo"
+company: "Pythian"
 date: "December 22, 2015"
@@
                           submitButton("Stats!"),
-                          # uiOutput('tableStat'),
-                          uiOutput('table')
+                          # uiOutput('tableResults'),
+                          uiOutput('table'),
+                          uiOutput('tableResults')
                           )
@@
 
-                #DataStat = reactive({
-                #  sql <- paste()
-                #  
-                #  rs <- dbGetQuery(con, sql)
-                #  return(rs)
-                #  
-                #})
+                DataResults = reactive({
+                   f<-function(x) {as.character(input$filtering)}
+                  sql <- paste("SELECT * FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 10", sep = "")
 
-                #output$tableStat <- renderTable({DataStat()})
+                  rs <- dbGetQuery(con, sql)
+                  return(rs)
+
+                })
+
+                output$tableResults <- renderTable({DataResults()})
 
@@
 - Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
+- If you want to have the results ordered by the last FTS expression in the WHERE clause, don't use `ORDER BY`, as MySQL
+  forcely does an order internally. 
+- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
+  it consumes a considerably amount of CPU.
 

### 2015-12-28 article.rmd
publishing for testing

@@
 title: "InnoDB Full text Search hacking with Shiny/R"
-author: "Emanuel Calvo"
-company: "Pythian"
+author: "Emanuel Calvo - Pythian"
 date: "December 22, 2015"
@@
 #knitr::opts_chunk$set(width=9999)
+# opts_chunk$set(fig.width=9999) #, fig.height=6))
 
@@
 
+## Version
+
+ **BETA - NOT FINAL**
+
 This R Markdown document is made interactive using Shiny.
@@
 
+Repository available at [Github](https://github.com/3manuek/fts_article).
+
+## Credits
+
+Author: Emanuel Calvo
+Company: Pythian
+
+Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
+
+## Requirements
+
 > You only need to have a MySQL sandbox with the 5.7 version.
 
-## Moto of the article
+## Some initial thoughts
 
@@
 
-## Stats
+## Considerations if you are planning to use Innodb's Full text search
+
+- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
+  size is crucial when using this.
+- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
+  single fields. 
+- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the 
+  FTS functions to avoid double executions. You can do this using somehting like:
+  `match(content,title) against ("first (<second >third)") > 1 `
+  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches. 
+- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
+  maintance on the table.
+- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
+- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as MySQL does the order forcelly internally. Also,
+  the behavior is different from the default as it returns the heaviest at the top (like ORDER BY rank DESC).
+- If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
+- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
+  it consumes a considerably amount of CPU. MySQL orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
+- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like count(*), will use
+  the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.
+
+```
+(test) >  explain  select  * 
+                      from bookContentByLine 
+                      where match(content) against ("+home" IN BOOLEAN MODE) 
+                      ORDER BY FTS_DOC_ID  
+                      LIMIT 10\G
+  select_type: SIMPLE
+        table: bookContentByLine
+         type: fulltext
+        Extra: Using where; Ft_hints: no_ranking; Using filesort
+
+(test) >  explain  select  * 
+                      from bookContentByLine 
+                      where match(content) against ("+home" IN BOOLEAN MODE) 
+                      LIMIT 10\G
+        table: bookContentByLine
+         type: fulltext
+        Extra: Using where; Ft_hints: no_ranking, limit = 10
+
+(test) >  explain  select  count(content) 
+                    from bookContentByLine 
+                    where match(content) against ("+home" IN BOOLEAN MODE) \G
+        table: bookContentByLine
+         type: fulltext
+        Extra: Using where; Ft_hints: no_ranking
+
+```
+- If you plan to use FTS_DOC_ID column with AUTO_INCREMENT option, have in mind that there is a limitation regarding this.You must
+  declare a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data type is stricted as `bigint unsigned`. I.e:
+
+```
+CREATE TABLE test ( FTS_DOC_ID bigint unsigned  AUTO_INCREMENT, 
+                    mainPk bigint, other text, 
+                    PRIMARY KEY(mainPk),
+                    UNIQUE(FTS_DOC_ID)
+                    );
+```
+
+## Stats of the datasets
+
+Here we prompt some stats regarding the used datasets for the tests and data written here:
 
@@
                    f<-function(x) {as.character(input$filtering)}
-                  sql <- paste("SELECT * FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 10", sep = "")
+                  sql <- paste("SELECT * FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") ", sep = "")
 
-                  rs <- dbGetQuery(con, sql)
+                  rs <- head(dbGetQuery(con, sql), 50)
                   return(rs)
@@
 
-## Recommendations
-
-- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
-  size is crucial when using this.
-- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
-  single fields. 
-- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the 
-  FTS functions to avoid double executions. You can do this using somehting like:
-  `match(content,title) against ("first (<second >third)") > 1 `
-  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches. 
-- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
-  maintance on the table.
-- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
-- If you want to have the results ordered by the last FTS expression in the WHERE clause, don't use `ORDER BY`, as MySQL
-  forcely does an order internally. 
-- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
-  it consumes a considerably amount of CPU.
-
----------------------- FROM HERE ONLY TRASH TO BE REMOVED -----------------------------
-
-Template
-
-```{r}
-inputPanel(
-  selectInput("n_breaks", label = "Number of bins:",
-              choices = c(10, 20, 35, 50), selected = 20),
-
-  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
-              min = 0.2, max = 2, value = 1, step = 0.2)
-)
-
-renderPlot({
-  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
-       xlab = "Duration (minutes)", main = "Geyser eruption duration")
-
-  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
-  lines(dens, col = "blue")
-})
-```
-
-## Embedded Application
-
-It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:
-
-```{r, echo=FALSE}
-shinyAppDir(
-  system.file("examples/06_tabsets", package="shiny"),
-  options=list(
-    width="100%", height=550
-  )
-)
-```
-
-Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.
-
-You can also use the `shinyApp` function to define an application inline rather then in an external directory.
-
-In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.
-
 ### Fine tuning
@@
 [Fine tuning](https://dev.mysql.com/doc/refman/5.7/en/fulltext-fine-tuning.html)
+
+### Features introduces
 

### 2015-12-28 load.R

@@ -56,6 +56,12 @@ dbGetQuery(con, "CREATE TABLE IF NOT EXISTS bookContentByLine
                    content text
                  ); ")
 
+# Other way to create the table within other PK:
+# node1 [localhost] {msandbox} (test) > CREATE TABLE test (  FTS_DOC_ID bigint unsigned  AUTO_INCREMENT, kk bigint, other text, PRIMARY KEY(kk),UNIQUE(FTS_DOC_ID));
+# Query OK, 0 rows affected (0,08 sec)
+
+
+
 #vector of ids for book downloading
 #randids <- abs(round(rnorm(opts$books)*1000))
 

### 2015-12-28 article.rmd
-
Changed the string connection.

@@
 # You do not want clear passwords, but is a sandbox.
+#con <- dbConnect(MySQL(), 
+#                 user="msandbox", password="msandbox", port=14901 
+#                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock", 
+#                 host="127.0.0.1")
+
 con <- dbConnect(MySQL(), 
-                 user="msandbox", password="msandbox", port=14901 
-                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock", 
-                 host="127.0.0.1")
+                 user="ftsinnodb", password="elfilete", port=3306 
+                 ,dbname="ftsinnodb", #socket="/tmp/mysql_sandbox14901.sock", 
+                 host="db4free.net")
 
@@
 Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
+
+Thanks db4free.net for the MySQL free online database!
 
@@
 - If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
+- Create the FULLTEXT index after the data is loaded [InnoDB bulk load](http://dev.mysql.com/doc/refman/5.7/en/optimizing-innodb-bulk-data-loading.html).
+When restoring FTS backups, you will probably hit the "ERROR 182 (HY000) at line nn: Invalid InnoDB FTS Doc ID".
 - Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and

### 2015-12-28 article.rmd
change on the string conn
Added comments
Clean some testing code

@@
 
-#Comment if you want to avoid reload all the data
+# Comment if you want to avoid reload all the data
+# If you want to start from scratch, uncomment the following line.
+# Otherwise, you can restore the dumps located in the dumps folder
 #source('~/fts_article/load.R')
@@
                    f<-function(x) {as.character(input$filtering)}
-                  sql <- paste("SELECT * FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") ", sep = "")
+                  sql <- paste("SELECT FTS_DOC_ID, bookid, 
+                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,"), content 
+                              FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 50", sep = "")
 
-                  rs <- head(dbGetQuery(con, sql), 50)
+                  rs <- dbGetQuery(con, sql)
                   return(rs)

### 2015-12-29 article.rmd
comments

@@
 ```{r init, message = FALSE, echo=FALSE}
+library(png)
 #options(width = 9999)
-#knitr::opts_chunk$set(width=9999)
+knitr::opts_chunk$set(fig.width=9999)
 # opts_chunk$set(fig.width=9999) #, fig.height=6))

### 2015-12-31 article.rmd
Pushed more details and contents

--- 2015-12-28 article.rmd
@@
 ```{r init, message = FALSE, echo=FALSE}
+library(png)
 #options(width = 9999)
-#knitr::opts_chunk$set(width=9999)
+#knitr::opts_chunk$set(fig.width=9999)
 # opts_chunk$set(fig.width=9999) #, fig.height=6))
@@
 
-This R Markdown document is made interactive using Shiny.
-Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately.
+This R Markdown document is made interactive using Shiny. Unlike the more traditional
+workflow of creating static reports, you can now create documents that allow your
+readers to change the assumptions underlying your analysis and see the results immediately.  
 To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).
@@
 
-> You only need to have a MySQL sandbox with the 5.7 version.
+If you want to run this locally, please install a MySQL 5.7 instance and:
+
+- Uncomment the line __source('~/fts_article/load.R')__ and run the report, or
+- Restore the dumps located in the `dumps` folder. Uncomment the table creation.
 
@@
 
-A couple of days ago one of our customers came up with a question regarding FTS over InnoDB engine. Althuogh the question
-is not entirely related with most of the content written here, I came up with the conclusion that FTS sometimes is a misunderstood feature.
+A couple of days ago one of our customers came up with a question regarding FTS
+over InnoDB engine. Although the question is not answered in the current article,
+I came up with the conclusion that FTS is sometimes misunderstood.
 
-The point of this article will be clarify when each algorithm fits in each situation.
+The point of this "article" is to show dynamically how the search algorithms work,
+using non-fictional data (data sources were downloaded from www.gutenberg.org
+project) within an easy interface.
 
-Both official documentation and articles are propagated over internet, however sometimes most of them lacks to answer the
-main questions and test cases. In order to follow a reproducible article, I used public available data sources (www.gutenberg.org).
+In order to show the effects off the field sizes over the query expansion algorithm,
+you will see two main tables (bookContent and bookContentByLine) both containing
+the same books in different approaches: by line and by paragraph. You'll see the
+noise generated by the `QUERY EXPANSION` algorithm when phrases are too large.
 
-In order to show the effects off the field sizes on the ranking algorithms, you will see two main tables (bookContent and 
-bookContentByLine) which contains the books parsed by paragraph and by line respectively. You'll see the noise generated
-by the `QUERY EXPANSION` algorithm when phrases are too large, and that's why you see both tables.
+The current article has been developed using Shiny/R in order to allow you to see
+the effects of the algorithms. It connects to a open database at `db4free.net`,
+which unfortunately is not 100% available.
 
-The current article has been developed using Shiny/R in order to allow you to see the effects of the algorithms.
+## Action!
 
 Here is an example of how ranks differ among algorithms and field sizes using the word 'country':
+
+To start showing the effects of the algorithms, the following example searches
+the word 'country' using `query expansion`. This means that we are not looking
+only the exact matches, but also the entries that appear the most when the
+the exact match has been found.
+
+In the `SELECT` clause you'll see both FTS expressions using `NATURAL LANGUAGE`
+with query expansion and `BOOLEAN`modes respectively.
 
@@
 
-SELECT  content, group_concat(it.POSITION) as positions, 
+SELECT  content, bookid, group_concat(it.POSITION) as pos,
         round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank,
-        round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, 
-        length(content) as len 
+        round(MATCH(content) AGAINST ("country" IN BOOLEAN MODE)) as BoolRank
       FROM bookContentByLine bl join information_schema.INNODB_FT_INDEX_TABLE it
@@
       WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
-          AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY 3 DESC LIMIT 10 ;
+          AND it.WORD = 'country'
+     GROUP BY FTS_DOC_ID
+     ORDER BY 4 DESC
+     LIMIT 10 ;
 
-+---------------------------------------------------------------------------+-----------+--------+-------------+------+
-| content                                                                   | positions | QERank | BooleanRank | len  |
-+---------------------------------------------------------------------------+-----------+--------+-------------+------+
-| "country may be ennobled, and under its auspices may be verified that"    | 1         |     54 |           4 |   70 |
-| "country, under a most excellent president, wherein all cities had their" | 1         |     53 |           4 |   73 |
-| "Again, the prince who holds a country differing in the above respects"   | 31        |     51 |           4 |   71 |
-| "country districts. As Lucca had five gates, he divided his own country"  | 1,63      |     51 |           8 |   72 |
-| "inflicting immense damage upon the country. Whereupon the Florentines"   | 36        |     50 |           4 |   71 |
-| "The cities of Germany are absolutely free, they own but little country"  | 64        |     49 |           4 |   72 |
-| "citizens of Fermo, to whom the slavery of their country was dearer than" | 49        |     48 |           4 |   73 |
-| "is useful in two ways. Firstly, he learns to know his country, and"      | 55        |     46 |           4 |   68 |
-| "agree that he should retain any lordship over the country. Because the"  | 51        |     45 |           4 |   72 |
-| "his own parts of the country, according to the authority he had assumed" | 22        |     43 |           4 |   73 |
-+---------------------------------------------------------------------------+-----------+--------+-------------+------+
++-------------------------------------------------------------------------------+--------+------+--------+----------+
+| content                                                                       | bookid | pos  | QERank | BoolRank |
++-------------------------------------------------------------------------------+--------+------+--------+----------+
+| "country in September, migrating into Holland, and leave their mates behind"  |  15707 | 1    |    105 |        7 |
+| "unsatisfied desire to serve his country, the two prevalent enthusiasms at"   |  15707 | 33   |     98 |        7 |
+| "Language, Vol. I. p. 212. In this country, where four or five horses travel" |  15707 | 35   |     93 |        7 |
+| "inflicting immense damage upon the country. Whereupon the Florentines"       |   1232 | 36   |     89 |        7 |
+| "made for a country of twenty or thirty millions’ population, can be laid"  |  39064 | 12   |     89 |        7 |
+| "The spiders of this country manufacture nets of various forms, adapted to"   |  15707 | 21   |     87 |        7 |
+| "a velvet-covered arm-chair at my head! This country is too decadent"         |  33415 | 45   |     86 |        7 |
+| "country may be ennobled, and under its auspices may be verified that"        |   1232 | 1    |     84 |        7 |
+| "name. The writer of this unpublished pamphlet sees his country in a"         |  39064 | 56   |     84 |        7 |
+| "In our country, Mr. Pennant informs us, that some quails migrate, and"       |  15707 | 8    |     83 |        7 |
+| "all the morning in passing over the adjacent country." (Voyage to Senegal,"  |  15707 | 46   |     82 |        7 |
+| "the electoral system of the country. Immediately an outcry burst out"        |  39064 | 29   |     82 |        7 |
+| "country, under a most excellent president, wherein all cities had their"     |   1232 | 1    |     81 |        7 |
+| "Though in this country horses shew little vestiges of policy, yet in the"    |  15707 | 16   |     81 |        7 |
+| "country districts. As Lucca had five gates, he divided his own country"      |   1232 | 1,63 |     80 |       14 |
++-------------------------------------------------------------------------------+--------+------+--------+----------+
+15 rows in set (1,16 sec)
 
@@
 ```
-| "country districts. As Lucca had five gates, he divided his own country"  | 1,63      |     51 |           8 |   72 |
+| "country districts. As Lucca had five gates, he divided his own country" |   1232 | 1,63 |     80 |       14 |
 ```
@@
 
-(test) > SELECT FTS_DOC_ID, 
+(test) > SELECT bookid, FTS_DOC_ID,
                 group_concat(it.POSITION) as positions,
                 round(MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)) as QERank,
-                round(MATCH(content) AGAINST ("country*" IN BOOLEAN MODE)) as BooleanRank, 
+                round(MATCH(content) AGAINST ("country" IN BOOLEAN MODE)) as BooleanRank,
                 length(content) as len
@@
           WHERE  MATCH(content) AGAINST ("country" IN NATURAL LANGUAGE MODE WITH QUERY EXPANSION)
-              AND lower(it.WORD) like '%country%' group by FTS_DOC_ID ORDER BY QERank DESC LIMIT 10 ;
-+------------+------------------+--------+-------------+------+
-| FTS_DOC_ID | positions        | QERank | BooleanRank | len  |
-+------------+------------------+--------+-------------+------+
-|        757 | 6858             |   2804 |           1 | 8101 | <<< This one on the top does only 1 exact match.
-|        735 | 2842             |    925 |           1 | 3133 |
-|        157 | 747              |    698 |           1 | 2385 |
-|        733 | 265,1342,816,750 |    569 |           2 | 1765 |
-|        739 | 1171             |    540 |           1 | 1659 |
-|        249 | 197,739          |    435 |           1 | 1368 |
-|        279 | 664              |    428 |           1 | 1365 |
-|        411 | 468,410          |    396 |           2 | 1170 |
-|        155 | 31,495,158,525   |    363 |           5 | 1378 |  <<< This doc contains 5 exact matches!
-|        327 | 60               |    362 |           1 | 1372 |
-+------------+------------------+--------+-------------+------+
-10 rows in set (4,09 sec)
+              AND it.WORD = 'country'
+          GROUP BY FTS_DOC_ID
+          ORDER BY QERank DESC
+          LIMIT 10 ;
+
++--------+------------+-----------------+--------+-------------+-------+
+| bookid | FTS_DOC_ID | positions       | QERank | BooleanRank | len   |
++--------+------------+-----------------+--------+-------------+-------+
+|  16452 |      17637 | 942,2552,9084   |  32494 |          10 | 51790 |
+|  16452 |      17827 | 31699           |  30232 |           3 | 51701 |
+|  16452 |      17761 | 667,47646       |  29517 |           7 | 50264 |
+|  16452 |      17791 | 13566           |  28888 |           3 | 49129 |
+|  16452 |      17927 | 23259,7044      |  26731 |           7 | 48983 |
+|  16452 |      17839 | 9012,199        |  24933 |           7 | 44451 |
+|  16452 |      17815 | 29318           |  24745 |           3 | 44011 |
+|  16452 |      17729 | 895,16485,24034 |  23305 |          10 | 42612 |
+|  16452 |      17621 | 1765            |  19935 |           3 | 36698 |
+|  16452 |      17803 | 3942            |  17552 |           3 | 30586 |
++--------+------------+-----------------+--------+-------------+-------+
+10 rows in set (1,88 sec)
 ```
-
-This kind of searches is useful if you want to find entries with words that appear frequently within your search.
 
@@
 
-Following stats are from the bookContent table (which has an average length of `r rs$len` characters ).
+The query expansion is useful when you intend to search which entries contain
+more words that appear frequently within the search term. Having large text fields
+increase the probability to have more words that appear among the search term.
+In the case of `bookContent` table (by paragraph table), the average field size
+is `r rs$len` characters.
+
+## The INNODB_FT_INDEX_TABLE
+
+There is a way to play with the contents of the FTS indexes. As you ma noticed
+in the previos examples, I used the `set global innodb_ft_aux_table = 'test/bookContent';`
+statement, which loads the index content to memory for an easy querying.
+
+i.e. You can easily get the most frequent tokens:
+
+```
+(test) > select WORD,count(*)
+           FROM information_schema.INNODB_FT_INDEX_TABLE   
+           group by WORD having count(*) > 1000
+           order by 2
+           limit 10;
++--------+----------+
+| WORD   | count(*) |
++--------+----------+
+| should |     1023 |
+| yet    |     1027 |
+| any    |     1070 |
+| like   |     1071 |
+| been   |     1073 |
+| first  |     1080 |
+| nor    |     1087 |
+| your   |     1106 |
+| thou   |     1130 |
+| shall  |     1164 |
++--------+----------+
+10 rows in set (5,40 sec)
+```
+
+Probably it isn't a very useful information as most of this words appears too frequently
+and are modal verbs,  adverbs,  pronouns, determiners, etc.
+
+It could be the case that you are not interested on indexing those words. If that's
+the case you can add them as `stopwords` in your own stopwords table. I built I
+query for this situation to allow us to build the stopwords table using the current
+words that we want to add to the filtering:
+
+```
+(test) > select group_concat(WORD) FROM (select distinct WORD  
+  FROM information_schema.INNODB_FT_INDEX_TABLE               
+  group by WORD having count(*) > 1000) d\G
+*************************** 1. row ***************************
+group_concat(WORD): all,and,any,been,but,can,first,had,has,have,her,him,his,into,
+its,like,may,more,nor,not,now,one,only,other,our,said,shall,she,should,some,such,
+than,their,them,then,there,these,they,those,thou,thus,thy,time,were,which,would,
+yet,you,your
+1 row in set (5,28 sec)
+
+```
+
+Let's build our filter table using both default and new entries and keeping the
+alphabetical order:
+
+```
+CREATE TABLE bookContentByLine_stopwords(value VARCHAR(30)) ENGINE = INNODB;
+
+INSERT INTO bookContentByLine_stopwords
+SELECT value FROM (
+    SELECT value FROM
+    INFORMATION_SCHEMA.INNODB_FT_DEFAULT_STOPWORD
+    UNION
+    SELECT DISTINCT WORD as value  
+      FROM information_schema.INNODB_FT_INDEX_TABLE               
+      GROUP BY WORD having count(*) > 1000
+) allEntries
+ORDER BY value ASC;
+
+DROP INDEX ftscontent ON bookContentByLine;
+SET GLOBAL innodb_ft_server_stopword_table = 'test/bookContentByLine_stopwords';
+CREATE FULLTEXT INDEX ftscontent ON bookContentByLine(content);
+
+```
+
+TODO: Check index sizes
 
@@
           from information_schema.INNODB_FT_INDEX_TABLE
-          WHERE DOC_ID = 733 AND lower(WORD) like '%country%';
+          WHERE WORD = 'country';
 +------------------+--------------+-------------+-----------+--------+----------+
@@
   it consumes a considerably amount of CPU. MySQL orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
-- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like count(*), will use
+- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like `count(*)`, will use
   the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.

### 2016-01-01 article.rmd
Added more details.

@@
 which unfortunately is not 100% available.
+
+For the sake of simplicity, in this article we won't go through the FTS parsers.
+Probably that would be material for a future post.
 
@@
 
-There is a way to play with the contents of the FTS indexes. As you ma noticed
-in the previos examples, I used the `set global innodb_ft_aux_table = 'test/bookContent';`
+There is a way to play with the contents of the FTS indexes. As you may noticed
+in the previous examples, I used the `set global innodb_ft_aux_table = 'test/bookContent';`
 statement, which loads the index content to memory for an easy querying.
@@
 ```
-(test) > select WORD,count(*)
+(test) > SELECT WORD,count(*)
            FROM information_schema.INNODB_FT_INDEX_TABLE   
@@
 Probably it isn't a very useful information as most of this words appears too frequently
-and are modal verbs,  adverbs,  pronouns, determiners, etc.
-
-It could be the case that you are not interested on indexing those words. If that's
-the case you can add them as `stopwords` in your own stopwords table. I built I
-query for this situation to allow us to build the stopwords table using the current
-words that we want to add to the filtering:
+and are modal verbs,  adverbs,  pronouns, determiners, etc. It could be the case that you
+are not interested on indexing those words. If that's the case you can add them as `stopwords` 
+in your own stopwords table. Specially if you are more interested in boolean searches, loosing
+some part of the language expressions. I built I query for this situation to allow us to build 
+the stopwords table using the current words that we want to add to the filtering:
 
@@
 
-## Considerations if you are planning to use Innodb's Full text search
+## Considerations and recommendations
 
@@
   size is crucial when using this.
-- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
-  single fields.
+- FTS is not the best fit for exact string matches in single columns. You don't want to use FTS for searching emails in
+  a single column, name and lastname fields , i.e. For those, you'll probably use other techniques as reverse searches or
+  exact match operator (=).
+- Keep your FTS indexes short. Do not add ALL the text columns. Parse first from your application the user search and 
+  adapt the query. 
 - If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the
-  FTS functions to avoid double executions. You can do this using somehting like:
+  FTS functions to avoid double executions. You can do this using something like:
   `match(content,title) against ("first (<second >third)") > 1 `

### 2016-01-01 article.rmd
fix table

@@
 
-+-------------------------------------------------------------------------------+--------+------+--------+----------+
-| content                                                                       | bookid | pos  | QERank | BoolRank |
-+-------------------------------------------------------------------------------+--------+------+--------+----------+
-| "country in September, migrating into Holland, and leave their mates behind"  |  15707 | 1    |    105 |        7 |
-| "unsatisfied desire to serve his country, the two prevalent enthusiasms at"   |  15707 | 33   |     98 |        7 |
-| "Language, Vol. I. p. 212. In this country, where four or five horses travel" |  15707 | 35   |     93 |        7 |
-| "inflicting immense damage upon the country. Whereupon the Florentines"       |   1232 | 36   |     89 |        7 |
-| "made for a country of twenty or thirty millions’ population, can be laid"  |  39064 | 12   |     89 |        7 |
-| "The spiders of this country manufacture nets of various forms, adapted to"   |  15707 | 21   |     87 |        7 |
-| "a velvet-covered arm-chair at my head! This country is too decadent"         |  33415 | 45   |     86 |        7 |
-| "country may be ennobled, and under its auspices may be verified that"        |   1232 | 1    |     84 |        7 |
-| "name. The writer of this unpublished pamphlet sees his country in a"         |  39064 | 56   |     84 |        7 |
-| "In our country, Mr. Pennant informs us, that some quails migrate, and"       |  15707 | 8    |     83 |        7 |
-| "all the morning in passing over the adjacent country." (Voyage to Senegal,"  |  15707 | 46   |     82 |        7 |
-| "the electoral system of the country. Immediately an outcry burst out"        |  39064 | 29   |     82 |        7 |
-| "country, under a most excellent president, wherein all cities had their"     |   1232 | 1    |     81 |        7 |
-| "Though in this country horses shew little vestiges of policy, yet in the"    |  15707 | 16   |     81 |        7 |
-| "country districts. As Lucca had five gates, he divided his own country"      |   1232 | 1,63 |     80 |       14 |
-+-------------------------------------------------------------------------------+--------+------+--------+----------+
++-----------------------------------------------------------------------------+--------+-----+--------+--------+
+|content                                                                      | bookid | pos | QERank |BoolRank|
++-----------------------------------------------------------------------------+--------+-----+--------+--------+
+|"country in September, migrating into Holland, and leave their mates behind" |  15707 | 1   |    105 |      7 |
+|"unsatisfied desire to serve his country, the two prevalent enthusiasms at"  |  15707 | 33  |     98 |      7 |
+|"Language, Vol. I. p. 212. In this country, where four or five horses travel"|  15707 | 35  |     93 |      7 |
+|"inflicting immense damage upon the country. Whereupon the Florentines"      |   1232 | 36  |     89 |      7 |
+|"made for a country of twenty or thirty millions’ population, can be laid" |  39064 | 12  |     89 |      7 |
+|"The spiders of this country manufacture nets of various forms, adapted to"  |  15707 | 21  |     87 |      7 |
+|"a velvet-covered arm-chair at my head! This country is too decadent"        |  33415 | 45  |     86 |      7 |
+|"country may be ennobled, and under its auspices may be verified that"       |   1232 | 1   |     84 |      7 |
+|"name. The writer of this unpublished pamphlet sees his country in a"        |  39064 | 56  |     84 |      7 |
+|"In our country, Mr. Pennant informs us, that some quails migrate, and"      |  15707 | 8   |     83 |      7 |
+|"all the morning in passing over the adjacent country." (Voyage to Senegal," |  15707 | 46  |     82 |      7 |
+|"the electoral system of the country. Immediately an outcry burst out"       |  39064 | 29  |     82 |      7 |
+|"country, under a most excellent president, wherein all cities had their"    |   1232 | 1   |     81 |      7 |
+|"Though in this country horses shew little vestiges of policy, yet in the"   |  15707 | 16  |     81 |      7 |
+|"country districts. As Lucca had five gates, he divided his own country"     |   1232 | 1,63|     80 |     14 |
++-----------------------------------------------------------------------------+--------+-----+--------+--------+
 15 rows in set (1,16 sec)

### 2016-01-04 article.rmd
Some cosmetic changes before push.

@@
 ---
-title: "InnoDB Full text Search hacking with Shiny/R"
+title: "InnoDB's Full Text Search Dynamic article with Shiny/R"
 author: "Emanuel Calvo - Pythian"
-date: "December 22, 2015"
+date: "January 3rd, 2016"
 output: html_document
@@
   maintance on the table.
-- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
-- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as MySQL does the order forcelly internally. Also,
-  the behavior is different from the default as it returns the heaviest at the top (like ORDER BY rank DESC).
+- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE. This affects the ranking score (try __+bad (<feeling >thing)__ i.e.)
+- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as InnoDB does the order after retrieve the doc ids .
+Also,the behavior is different from the default as it returns the heaviest at the top (like an __ORDER BY rank DESC__).
 - If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
@@
 When restoring FTS backups, you will probably hit the "ERROR 182 (HY000) at line nn: Invalid InnoDB FTS Doc ID".
-- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
-  it consumes a considerably amount of CPU. MySQL orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
+- Try to avoid using use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
+  it consumes a considerably amount of CPU. InnoDB orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
 - Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like `count(*)`, will use
@@
                   sql <- paste("SELECT FTS_DOC_ID, bookid,
-                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,"), content
+                              MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") as rankScore, content
                               FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 50", sep = "")
@@
                                       max(scoreQE) as maxQE,
-                                      -- max(scoreNL) as maxNL,
+                                      max(scoreNL) as maxNL,
                                       avg(scoreBoolean) as avgBoolean,
                                       avg(scoreQE) as avgQE,
-                                      -- avg(scoreNL) as avgNL,
+                                      avg(scoreNL) as avgNL,
                                       avg(lengthContent) avgFieldLength,

### 2016-02-02 article.rmd
CHanged to work locally as db4free died.

@@
 # You do not want clear passwords, but is a sandbox.
+con <- dbConnect(MySQL(),
+                 user="msandbox", password="msandbox", port=13454
+                 ,dbname="test", socket="/tmp/mysql_sandbox13454.sock",
+                 host="127.0.0.1")
+
 #con <- dbConnect(MySQL(),
-#                 user="msandbox", password="msandbox", port=14901
-#                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock",
-#                 host="127.0.0.1")
-
-con <- dbConnect(MySQL(),
-                 user="ftsinnodb", password="elfilete", port=3306
-                 ,dbname="ftsinnodb", #socket="/tmp/mysql_sandbox14901.sock",
-                 host="db4free.net")
+#                 user="ftsinnodb", password="elfilete", port=3306
+#                 ,dbname="ftsinnodb", #socket="/tmp/mysql_sandbox14901.sock",
+#                 host="db4free.net")
 
@@
 
+### Bug
+
+```
+emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
+ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
+```
+
+It works if loaded ddl and data separately.
+
+```
+emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < ddl.dump 
+emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < onlyData.dump 
+emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
+ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
+```
+
+Changed innodb_autoinc_lock_mode to 0 (old behavior) and 2, with the same results.
+Looks like FTS in InnoDB does not like metadata locks (as we are using a single transaction
+to load the data).
+
 ### Fine tuning
@@
 [Fine tuning](https://dev.mysql.com/doc/refman/5.7/en/fulltext-fine-tuning.html)
+[Performance](https://blogs.oracle.com/mysqlinnodb/entry/innodb_full_text_search_performance)
 

### 2016-02-02 load.R

@@ -32,8 +32,8 @@ for(con in all_cons) dbDisconnect(con)
 
 # You do not want clear passwords, but is a sandbox.
 con <- dbConnect(MySQL(), 
-                 user="msandbox", password="msandbox", port=14901 
-                 ,dbname="test", socket="/tmp/mysql_sandbox14901.sock", 
+                 user="msandbox", password="msandbox", port=13454 
+                 ,dbname="test", socket="/tmp/mysql_sandbox13454.sock", 
                  host="127.0.0.1")
 # on.exit(dbDisconnect(con))
 

### 2016-02-12 article.rmd
https://bugs.mysql.com/bug.php?id=80347 commited bug report, added more information
started abuot zipf law

@@
 
+### Going ahead on choosing stop words
+
+The full [article](https://en.wikipedia.org/wiki/Zipf's_law) is amazingly interesting. In a brief, it says that
+the most frequent word will occur approximately twice as often as the second most frequent word, three times as 
+often as the third most frequent word, and so on (rank-frequency distribution is an inverse relation).
+
+The idea here is to measure how much index do we safe cutting those words that are extremely frequent and don't add
+a necessary meaning to the search.
+
 ## Considerations and recommendations
@@
 ```
-- If you plan to use FTS_DOC_ID column with AUTO_INCREMENT option, have in mind that there is a limitation regarding this.You must
+- If you plan to use `FTS_DOC_ID` column with AUTO_INCREMENT option, have in mind that there is a limitation regarding this.You must
   declare a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data type is stricted as `bigint unsigned`. I.e:
@@
 
+### FT_QUERY_EXPANSION_LIMIT
+
+This variable controls the number of top matches when using `WITH QUERY EXPANSION` (affects only MyISAM).
+
+[reference](http://dev.mysql.com/doc/refman/5.7/en/server-system-variables.html#sysvar_ft_query_expansion_limit)
+
 ### Bug
+
+Submitted https://bugs.mysql.com/bug.php?id=80347
 

### 2016-04-19 article.rmd
Added files and compatibility with ansible

@@
 ---
-title: "InnoDB's Full Text Search Dynamic article with Shiny/R"
+title: "InnoDB's Full Text Search overview (with Shiny/R)"
 author: "Emanuel Calvo - Pythian"
@@
 
- **BETA - NOT FINAL**
-
 This R Markdown document is made interactive using Shiny. Unlike the more traditional
@@
 Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
-
-Thanks db4free.net for the MySQL free online database!
 

### 2016-04-19 load.R
Fixes

@@ -1,3 +1,4 @@
+#!/usr/bin/Rscript
 require(RMySQL)
 require(utils)
 library(stringr)

### 2016-04-20 article.rmd
Final changes

@@
 
-```{r init, message = FALSE, echo=FALSE}
+```{r init, message = FALSE, echo=FALSE,include=FALSE}
 library(png)
@@
 library(shiny)
-require(RMySQL)
+suppressWarnings(library(RMySQL))
 library(ggplot2)
@@
 # Otherwise, you can restore the dumps located in the dumps folder
+
 #source('~/fts_article/load.R')
+
+mypass <- scan(file = '.key', what = (pass=""))
+RDShost <- scan(file = '.RDShost', what = (RDShost=""))
 
@@
 con <- dbConnect(MySQL(),
-                 user="msandbox", password="msandbox", port=13454
-                 ,dbname="test", socket="/tmp/mysql_sandbox13454.sock",
-                 host="127.0.0.1")
-
-#con <- dbConnect(MySQL(),
-#                 user="ftsinnodb", password="elfilete", port=3306
-#                 ,dbname="ftsinnodb", #socket="/tmp/mysql_sandbox14901.sock",
-#                 host="db4free.net")
+                 user="mysql57", password=mypass, port=3306
+                 ,dbname="ftslab", #socket="/tmp/mysql_sandbox13454.sock",
+                 host=RDShost)
 
 ```
-
-## Version
-
-This R Markdown document is made interactive using Shiny. Unlike the more traditional
-workflow of creating static reports, you can now create documents that allow your
-readers to change the assumptions underlying your analysis and see the results immediately.  
-To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).
-
-Repository available at [Github](https://github.com/3manuek/fts_article).
 
@@
 Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
+
+> Repository available at [Github](https://github.com/3manuek/fts_article).
 
@@
 
-The point of this "article" is to show dynamically how the search algorithms work,
+The point of this _article_ is to show dynamically how the search algorithms work,
 using non-fictional data (data sources were downloaded from www.gutenberg.org
@@
 The current article has been developed using Shiny/R in order to allow you to see
-the effects of the algorithms. It connects to a open database at `db4free.net`,
-which unfortunately is not 100% available.
+the effects of the algorithms.
 
@@
 Probably that would be material for a future post.
+
+## Why I consider FTS sometimes misunderstood?
+
+FTS is a technology that can be use for any purpose, not only simple searches.
+There is a myth that the FTS only should be placed on clusters for that purpose,
+which I agree. However, certain bussiness rules require comple searches, and 
+having such feature can be a win.
+
+RDBMS aren't a good place for massive amount of FTS queries, without using any
+of the join capabilities that they offer, or the ACID complaints.
+
+As I said above, FTS is totally acceptable in RDBMS, if you are using at least
+one RDBMS critical feature, required by your bussines model.
 
@@
 
-```{r, echo=FALSE}
+```{r, ignore=FALSE, echo=FALSE, message=FALSE}
 
@@
             GROUP BY bookid;"
-rs <- dbGetQuery(con, sql)
-return(rs)
+rs1 <- dbGetQuery(con, sql)
 
-bookContentStats <- rs
+bookContentStats <- rs1
 
@@
             GROUP BY bookid;"
-rs <- dbGetQuery(con, sql)
-return(rs)
+rs2 <- dbGetQuery(con, sql)
 
-bookContentByLineStats <- rs
+bookContentByLineStats <- rs2
+
+```
+
+```{r, echo=FALSE}
+return(rs1)
+return(rs2)
 
@@
 
-### Bug
+### Bug 80347 - Invalid InnoDB FTS Doc ID
 
-Submitted https://bugs.mysql.com/bug.php?id=80347
+> Submitted https://bugs.mysql.com/bug.php?id=80347
 
@@
 ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
-```
 
-It works if loaded ddl and data separately.
-
-```
 emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < ddl.dump 
@@
 
-Changed innodb_autoinc_lock_mode to 0 (old behavior) and 2, with the same results.
-Looks like FTS in InnoDB does not like metadata locks (as we are using a single transaction
-to load the data).
+mysqldump is not very clever if you use `FTS_DOC_ID`:
+
+```
+2016-02-13T22:11:53.125300Z 19 [ERROR] InnoDB: Doc ID 10002 is too big. Its difference with largest used Doc ID 1 cannot exceed or equal to 10000
+```
+
+It takes dumps without considering the restriction coded in 
+`innobase/row/row0mysql.cc`:
+
+```
+Difference between Doc IDs are restricted within
+4 bytes integer. See fts_get_encoded_len() 
+```
+
+The fix to this is backuping the table by chunks of 10000 documents.
 
@@
 [Writting FTS parser plugins](https://dev.mysql.com/doc/refman/5.7/en/writing-full-text-plugins.html)
+
+## Version
+
+This R Markdown document is made interactive using Shiny. Unlike the more traditional
+workflow of creating static reports, you can now create documents that allow your
+readers to change the assumptions underlying your analysis and see the results immediately.  
+To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

### 2016-04-20 article.rmd
Fixes

@@
                  user="mysql57", password=mypass, port=3306
-                 ,dbname="ftslab", #socket="/tmp/mysql_sandbox13454.sock",
+                 ,dbname="ftslab", 
                  host=RDShost)
@@
 
-Author: Emanuel Calvo
-Company: Pythian
-
-Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
-
+> Author: Emanuel Calvo
+> Company: Pythian
+>
+> Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.
+>
 > Repository available at [Github](https://github.com/3manuek/fts_article).
@@
 ```
-set global innodb_ft_aux_table = 'test/bookContentByLine';
+set global innodb_ft_aux_table = 'ftslab/bookContentByLine';
 
@@
 
-The noise generated by the query expansion is expected and described in the official documentation [here.](https://dev.mysql.com/doc/refman/5.7/en/fulltext-query-expansion.html)
+The noise generated by the query expansion is expected and described in the official documentation [here](https://dev.mysql.com/doc/refman/5.7/en/fulltext-query-expansion.html).
 
@@
 
-(test) > set global innodb_ft_aux_table = 'test/bookContent';
+(ftslab) > set global innodb_ft_aux_table = 'ftslab/bookContent';
 Query OK, 0 rows affected (0,00 sec)
 
-(test) > SELECT bookid, FTS_DOC_ID,
+(ftslab) > SELECT bookid, FTS_DOC_ID,
                 group_concat(it.POSITION) as positions,
@@
 
-## The INNODB_FT_INDEX_TABLE
+## The `INNODB_FT_INDEX_TABLE`
 
 There is a way to play with the contents of the FTS indexes. As you may noticed
-in the previous examples, I used the `set global innodb_ft_aux_table = 'test/bookContent';`
+in the previous examples, I used the `set global innodb_ft_aux_table = 'ftslab/bookContent';`
 statement, which loads the index content to memory for an easy querying.
+
+If you use RDS, the option `innodb_ft_aux_table` is not available as it is GLOBAL
+and require SUPER privileges.
 
@@
 ```
-(test) > SELECT WORD,count(*)
+(ftslab) > SELECT WORD,count(*)
            FROM information_schema.INNODB_FT_INDEX_TABLE   
@@
 ```
-(test) > select group_concat(WORD) FROM (select distinct WORD  
+(ftslab) > select group_concat(WORD) FROM (select distinct WORD  
   FROM information_schema.INNODB_FT_INDEX_TABLE               
@@
 DROP INDEX ftscontent ON bookContentByLine;
-SET GLOBAL innodb_ft_server_stopword_table = 'test/bookContentByLine_stopwords';
+SET GLOBAL innodb_ft_server_stopword_table = 'ftslab/bookContentByLine_stopwords';
 CREATE FULLTEXT INDEX ftscontent ON bookContentByLine(content);
@@
 
-TODO: Check index sizes
+Checking the contents of the index is easy as:
 
 ```
-(test) > select *
+(ftslab) > select *
           from information_schema.INNODB_FT_INDEX_TABLE
-          WHERE WORD = 'country';
+          WHERE lower(WORD) like '%country%';
 +------------------+--------------+-------------+-----------+--------+----------+
@@
 
-(test) > select *
+(ftslab) > select *
           from information_schema.INNODB_FT_INDEX_TABLE
@@
 ```
-(test) >  explain  select  *
+(ftslab) >  explain  select  *
                       from bookContentByLine
@@
 
-(test) >  explain  select  *
+(ftslab) >  explain  select  *
                       from bookContentByLine
@@
 
-(test) >  explain  select  count(content)
+(ftslab) >  explain  select  count(content)
                     from bookContentByLine
@@
 ```
-- If you plan to use `FTS_DOC_ID` column with AUTO_INCREMENT option, have in mind that there is a limitation regarding this.You must
-  declare a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data type is stricted as `bigint unsigned`. I.e:
+
+- If you plan to use `FTS_DOC_ID` column with `AUTO_INCREMENT` option, 
+  have in mind that there is a limitation regarding this. You must declare 
+  a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data 
+  type is stricted as `bigint unsigned`. i.e:
 
